### 2020-12-27 paper/sex_selection_spacing-ver5.tex
Edited abstract and added transition/introduction paragraphs

\noindent Following the arrival of prenatal sex-determination technologies in the mid-1980s, 
India has experienced an increasingly male-biased sex ratio at birth, presumably from sex-selective 
abortions. 
Abortions lengthen birth intervals, but we know little about how birth spacing has changed or the 
effects of these changes. 
This paper shows that the length of birth intervals has increased overall, but well-educated women 
with no sons had the most substantial lengthening---and the most male-biased sex ratios. 
Furthermore, most of the changes took place immediately after the introduction of sex-selective abortions. 
Consequently, some women without sons now have longer birth intervals than those with sons, reversing 
India's traditional spacing pattern. 
Less-educated women continue short birth spacing when they have no sons, with only limited evidence 
of male-biased sex ratios. 
Because of the rapid lengthening of birth intervals, period fertility rates substantially overestimated 
how fast cohort fertility fell. 
Moreover, predicted cohort fertility is still 10\%--20\% above the period fertility rate. 
If lengthening of birth intervals occurs from repeated abortions, the associated short pregnancy spacing 
may counteract any positive effects of longer birth spacing. 
There is, however, no evidence of this effect on infant mortality. 
Finally, judging from sex ratios, sex-selective abortion use is not declining.

@@

The empirical analysis has three parts. 
First, I document the changes in birth spacing over time and how the introduction of sex 
selection influenced birth spacing. 
Second, I compare period fertility estimates with predicted cohort fertility based on the 
model of birth spacing. 
Finally, I examine how birth intervals affect infant mortality and whether the apparent 
increase in sex selection influenced mortality.

This section presents the empirical model for my analysis of birth spacing. 
The standard approach in the birth spacing literature is to use proportional hazard
models with a single exit---the birth of a child.%
\footnote{
See \citet{Sheps1970} and \citet{Newman1984} for early discussions of why 
hazard models are the preferred way to deal with the censoring of birth
intervals.
}
There are two problems with the standard approach in this setting.

@@

@@

% Explain question and what I am doing
The first question I address is how birth spacing responded to the significant 
changes in India.
Figures \ref{fig:spacing_low_rural} through \ref{fig:spacing_highest_urban}
show the 25th, 50th, and 75th percentile birth interval durations in 
months, the sex ratio, and the probability of having a birth for each spell
by education levels and area of residence.%
\footnote{
Results for urban women without education, rural women with 12 or more years of 
education, and the fourth spell for women with 12 or more years of education are
in the online appendix because of relatively small samples.
}
The sex ratio graphs also show the natural sex ratio, approximately 51.2\% boys
\citep{ben-porath76b,jacobsen99,Portner2015b}.
The underlying values with bootstrapped standard errors and tests for statistically
significant differences across sex compositions are available in the online appendix.

>>>>

\noindent Following the arrival of prenatal sex-determination technologies in the mid-1980s, 
India has experienced an increasingly male-biased sex ratio at birth, presumably from sex-selective 
abortions. 
Abortions lengthen birth intervals, but we know little about how birth spacing has changed or the 
effects of these changes. 
This paper shows that, although the length of birth intervals increased overall, well-educated women 
with no sons had the most substantial lengthening---and the most male-biased sex ratios. 
Furthermore, most of these changes took place immediately after the introduction of prenatal 
sex-determination technologies. 
Consequently, some women without sons now have longer birth intervals than those with sons, reversing 
India's traditional spacing pattern. 
Less-educated women continue short birth spacing when they have no sons, with only limited evidence 
of male-biased sex ratios. 
Because of the rapid lengthening of birth intervals, period fertility rates substantially overestimated 
how fast cohort fertility fell. 
Moreover, predicted cohort fertility is still 10\%--20\% above the period fertility rate. 
If lengthening of birth intervals occurs from repeated abortions, the associated short pregnancy spacing 
may counteract any positive effects of longer birth spacing. 
There is, however, no evidence of this effect on infant mortality. 
Finally, judging from sex ratios, sex-selective abortion use is not declining.

@@

The empirical analysis has three parts. 
First, I document the changes in birth spacing over time and how the introduction of sex 
selection influenced birth spacing. 
Second, I compare period fertility estimates with predicted cohort fertility based on the 
model of birth spacing. 
Finally, I examine how birth intervals affect infant mortality and whether the apparent 
increase in sex selection influenced mortality. 
This section presents the empirical model for my analysis of birth spacing.

The standard approach in the birth spacing literature is to use proportional hazard
models with a single exit---the birth of a child---but there are two problems with the 
standard approach in this setting.%
\footnote{
See \citet{Sheps1970} and \citet{Newman1984} for early discussions of why 
hazard models are the preferred way to deal with the censoring of birth
intervals.
}

@@

I first discuss sample restrictions and their motivations, then introduce the potential 
issue of underreporting of female births and the imposed censoring of birth intervals.
Next, I discuss how the analyses are split by periods to capture the introduction
and changing legality in prenatal sex determination.
Finally, I introduce the explanatory variables, split by whether they enter as
proportional or nonproportional variables.

@@

% Explain question and what I am doing
The first question I address is how birth spacing responded to the significant 
changes in India.
I begin with a broad outline of how parity progression probabilities and
sex ratios have changed over time and by groups.
I then use these results to separately discuss the changes in birth spacing for 
women who do not appear to use sex selection and for those who do.

Figures \ref{fig:spacing_low_rural} through \ref{fig:spacing_highest_urban}
show the 25th, 50th, and 75th percentile birth interval durations in 
months, the sex ratio, and the probability of having a birth for each spell
by education levels and area of residence.%
\footnote{
Results for urban women without education, rural women with 12 or more years of 
education, and the fourth spell for women with 12 or more years of education are
in the online appendix because of relatively small samples.
}
The sex ratio graphs also show the natural sex ratio, approximately 51.2\% boys
\citep{ben-porath76b,jacobsen99,Portner2015b}.
The underlying values with bootstrapped standard errors and tests for statistically
significant differences across sex compositions are available in the online appendix.
